---
title: "CZK Reference Interest Rates"
author: "Jindra Lacko"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CZK_Reference_Interest_Rates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = T, fig.align="center", dpi = 150, out.width = "100%", fig.width = 8, fig.height = 4)
```


# PRIBOR - PRague InterBank Offered Rate

The PRIBOR rates are provided by ČNB as a courtesy of the Czech Financial Benchmark Facility s.r.o., the benchmark administrator of the PRIBOR benchmark.

The rates can be accessed for internal purposes free of charge via internet pages of the Czech National Bank - https://www.cnb.cz/en/financial-markets/money-market/pribor/format-of-the-pribor-rate-on-the-cnb-website/

The package `{czechrates}` provides a convenient way to access the information stored on the ČNB site from the comfort of your R session. It does not store the rates (this would be against the terms of use, and the data would get stale rather soon). As a consequence a working internet connection is required to use the package.

The package currently contains a single function: `czechrates::pribor()`. 

The `pribor()` function has two parameters:

- `date` = date valid for the PRIBOR rate; default is yesterday (`Sys.Date() - 1`)

- `maturity` = tenor of the PRIBOR rate; quoted as one of the standard maturities: 1D (overnight = default), 1W (weekly) and 1M, 2M, 3M, 6M, 9M and 1Y.

```{r static, message = F}
library(czechrates)

# relying on default for both date and tenor
pribor() 

# specific date, default tenor
pribor(as.Date("2020-05-20")) 

# three specific dates + a specific tenor (weekly rate)
three_dates <- as.Date(c("2020-05-11", "2020-05-12", "2020-05-13"))
three_dates

pribor(three_dates, "1W")

# four rates for a single date
pribor(as.Date("2020-05-20"), c("1D", "1W", "1M", "1Y")) 

```

Using `czechrates::pribor()` you can re-live some of the scariest phases of the Prague interbank market - such as the period around Asian crisis of 1997 (when overnight rates briefly went over 100% on annual basis).

```{r asian-fever}
library(czechrates)
library(ggplot2)
library(scales)

offered_rate <- pribor(seq(from = as.Date("1997-05-01"), 
                           to   = as.Date("1997-06-30"),
                           by = 1))

ggplot(data = offered_rate, aes(x = date_valid, y = PRIBOR_1D)) +
  geom_line(color = "red", size = 1.25) +
  scale_x_date(date_labels = "%d.%m.%Y") +
  scale_y_continuous(labels = percent_format()) +
  theme_bw() +
  labs(title = "A ghost of rates past...",
       y = "Overnight PRIBOR (per annum basis)") +
  theme(axis.title.x = element_blank())

```
To ilustrate the development of the main policy rate – i.e. `czechrates::repo2w()` – we can consider a more recent history:

```{r recent-development}
library(czechrates)
library(ggplot2)
library(scales)

policy_rate <- repo2w(seq(from = as.Date("2020-01-01"), 
                          to   = as.Date("2020-05-31"),
                          by = 1))

ggplot(data = policy_rate, aes(x = date_valid, y = REPO_2W)) +
  geom_line(color = "red", size = 1.25) +
  scale_x_date(date_labels = "%d.%m.%Y") +
  scale_y_continuous(labels = percent_format()) +
  theme_bw() +
  labs(title = "COVID-19 impact on CZK policy rate",
       y = "ČNB REPO rate (per annum basis)") +
  theme(axis.title.x = element_blank())

```
